---
title: "Poster Graphics"
output: html_notebook
---


Load Packages and Data

```{r}
## Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(pracma)
library(RColorBrewer)
library(EnhancedVolcano)
library(gridExtra)
library(pals)
library(stringr)
library(viridis)

memory.limit(size = 16000000)

## Integrated PDAC datasets 
load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC155698.integrated")

## BCCIO dataset
load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 123813\\BCCIOwithUMAP")

## Merged dataset (5000 cells from PDAC and BCC, random sample)
load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\BCC.PDAC.merge")

## Load functions
source("ClusteringAnalysisFunctions.R", echo=TRUE)
source("SeuratClusteringFunctions.R", echo=TRUE)

```



BCC Clustering UMAP

```{r}

BCCIO <- RenameIdents(object = BCCIO, 
                      'CD8_memory' = 'CD8+ memory T cells',
                      'CD4' = 'CD4+ T cells', 
                      'Tregs' = 'Regulatory T cells',
                      'CD8_activation' = 'CD8+ activated T cells',
                      'CD8_exhausted' = 'CD8+ exhausted T cells',
                      'Plasma' = 'Plasma cells',
                      'Tumor 1' = 'Malignant cells 1',
                      'Tumor 2' = 'Malignant cells 2',
                      'T_proliferation' = 'Proliferating T cells',
                      'NK' = 'NK cells',
                      'Endothelial' = 'Endothelial cells',
                      'DCs' = 'Dendritic cells',
                      'Misc.' = 'Other')

BCC.UMAP <- DimPlot(object = BCCIO, 
                    reduction ="umap", 
                    label = TRUE, 
                    pt.size = 0.3,
                    label.size = 6) + NoLegend()
BCC.UMAP

## Default dimensions, width = 1200px
## Saved as BCC_UMAP.jpeg

```



BCC UMAP separated by patient

```{r}

BCC.UMAP.patient <- DimPlot(object = BCCIO,
                            reduction = "umap",
                            label = FALSE,
                            split.by = c("patient"),
                            pt.size = 0.1,
                            ncol = 6)
BCC.UMAP.patient

## Dimensions: width = 1500px, height = 505px
## Post-processing to remove axes in powerpoint
## Saved as BCC_UMAP_Patient.JPG

```



BCC UMAP separated by responders vs. nonresponders

```{r}

BCC.UMAP.responder <- DimPlot(object = BCCIO,
                              reduction = "umap",
                              label = FALSE,
                              split.by = c("response"),
                              pt.size = 0.1,
                              ncol = 2)

BCC.UMAP.responder

## Dimensions: width = 1000px, height = 505px
## Saved as BCC_UMAP_Responder.jpeg

```



BCC UMAP separated by patient and treatment

```{r}

BCCIO$patient.treatment <- paste0(BCCIO$patient, BCCIO$treatment)

Idents(BCCIO) <- BCCIO$patient.treatment

BCC.patient001.pre.cells <- WhichCells(BCCIO, idents = c("su001pre"))
BCC.patient001.post.cells <- WhichCells(BCCIO, idents = c("su001post"))
BCC.patient002.pre.cells <- WhichCells(BCCIO, idents = c("su002pre"))
BCC.patient002.post.cells <- WhichCells(BCCIO, idents = c("su002post"))
BCC.patient003.pre.cells <- WhichCells(BCCIO, idents = c("su003pre"))
BCC.patient003.post.cells <- WhichCells(BCCIO, idents = c("su003post"))
BCC.patient004.pre.cells <- WhichCells(BCCIO, idents = c("su004pre"))
BCC.patient004.post.cells <- WhichCells(BCCIO, idents = c("su004post"))
BCC.patient005.pre.cells <- WhichCells(BCCIO, idents = c("su005pre"))
BCC.patient005.post.cells <- WhichCells(BCCIO, idents = c("su005post"))
BCC.patient006.pre.cells <- WhichCells(BCCIO, idents = c("su006pre"))
BCC.patient006.post.cells <- WhichCells(BCCIO, idents = c("su006post"))
BCC.patient007.pre.cells <- WhichCells(BCCIO, idents = c("su007pre"))
BCC.patient007.post.cells <- WhichCells(BCCIO, idents = c("su007post"))
BCC.patient008.pre.cells <- WhichCells(BCCIO, idents = c("su008pre"))
BCC.patient008.post.cells <- WhichCells(BCCIO, idents = c("su008post"))
BCC.patient009.pre.cells <- WhichCells(BCCIO, idents = c("su009pre"))
BCC.patient009.post.cells <- WhichCells(BCCIO, idents = c("su009post"))
BCC.patient010.pre.cells <- WhichCells(BCCIO, idents = c("su010pre"))
BCC.patient010.post.cells <- WhichCells(BCCIO, idents = c("su010post"))
BCC.patient012.pre.cells <- WhichCells(BCCIO, idents = c("su012pre"))
BCC.patient012.post.cells <- WhichCells(BCCIO, idents = c("su012post"))

BCC.UMAP.patient1.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient001.pre.cells, BCC.patient001.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su001", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient2.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient002.pre.cells, BCC.patient002.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su002", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient3.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient003.pre.cells, BCC.patient003.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su003", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient4.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient004.pre.cells, BCC.patient004.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su004", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient5.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient005.pre.cells, BCC.patient005.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su005", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient6.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient006.pre.cells, BCC.patient006.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su006", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient7.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient007.pre.cells, BCC.patient007.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su007", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient8.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient008.pre.cells, BCC.patient008.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su008", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient9.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient009.pre.cells, BCC.patient009.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su009", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient10.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient010.pre.cells, BCC.patient010.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su010", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))
BCC.UMAP.patient12.treatment <- DimPlot(object = BCCIO,
                                      reduction = "umap", 
                                      label = FALSE,
                                      group.by = c("patient.treatment"),
                                      cells.highlight = list(BCC.patient012.pre.cells, BCC.patient012.post.cells),
                                      cols.highlight = c("royalblue1", "darkslategray1"),
                                      cols = c("grey90"),
                                      sizes.highlight = 0.1,
                                      pt.size = 0.1) + NoLegend() + 
                                      labs(title = "su012", x = "",  y = "") + 
                                      theme(axis.line.y = element_blank(), axis.line.x = element_blank(),
                                            axis.text.y = element_blank(), axis.text.x = element_blank(),
                                            axis.ticks.y = element_blank(), axis.ticks.x = element_blank(),
                                            plot.title = element_text(hjust = 0.5))

BCC.UMAP.patient.treatment <- grid.arrange(BCC.UMAP.patient1.treatment, BCC.UMAP.patient2.treatment, BCC.UMAP.patient3.treatment, BCC.UMAP.patient4.treatment, BCC.UMAP.patient5.treatment, BCC.UMAP.patient6.treatment, BCC.UMAP.patient7.treatment, BCC.UMAP.patient8.treatment, BCC.UMAP.patient9.treatment, BCC.UMAP.patient10.treatment, BCC.UMAP.patient12.treatment,  nrow=2)

BCC.UMAP.patient.treatment

## Dimensions: width = 1500px, height = 500px
## Post-processing to add legend in Powerpoint
## Saved as: BCC_UMAP_Patient_Treatment.JPG

```



BCC Clustering Marker Genes

```{r}

BCC.cluster.order <- c('CD4+ T cells', 'CD8+ memory T cells', 'CD8+ activated T cells', 'CD8+ exhausted T cells', 'Proliferating T cells', 'Regulatory T cells', 'NK cells', 'B cells', 'Plasma cells', 'Macrophages', 'Dendritic cells', 'pDCs', 'Myofibroblasts', 'CAFs', 'Malignant cells 1', 'Malignant cells 2', 'Endothelial cells', 'Melanocytes', 'Other')

BCCIO <- RenameIdents(object = BCCIO, 
                      'CD4+ T cells' = 'CD4+ T cells', 
                      'CD8+ memory T cells' = 'CD8+ memory T cells',
                      'CD8+ activated T cells' = 'CD8+ activated T cells',
                      'CD8+ exhausted T cells' = 'CD8+ exhausted T cells',
                      'Proliferating T cells' = 'Proliferating T cells',
                      'Regulatory T cells' = 'Regulatory T cells',
                      'NK cells' = 'NK cells',
                      'B cells' = 'B cells',
                      'Plasma cells' = 'Plasma cells',
                      'Macrophages' = 'Macrophages',
                      'Dendritic cells' = 'Dendritic cells',
                      'pDCs' = 'pDCs', 
                      'Myofibroblasts' = 'Myofibroblasts',
                      'CAFs' = 'CAFs',
                      'Malignant cells 1' = 'Malignant cells 1',
                      'Malignant cells 2' = 'Malignant cells 2',
                      'Endothelial cells' = 'Endothelial cells',
                      'Melanocytes' = 'Melanocytes',
                      'Other' = 'Other')

features <- c('CD3D', 'CD3E', 'CD2', 'CD8A', 'GZMA', 'GZMB', 'CD4', 'FOXP3', 'KLRC1', 'KLRC3', 'NKG7', 'CD19', 'CD79A', 'SLAMF7', 'IGKC', 'FCGR2A', 'CSF1R', 'FLT3', 'CLEC4C', 'COL1A2', 'MCAM', 'MYLK', 'FAP', 'PDPN', 'EPCAM', 'TP63', 'PECAM1', 'VWF', 'PMEL', 'MLANA')

BCC.cluster.markers <- DotPlot(BCCIO, features = features, assay = 'RNA', cols = c('Azure', 'Blue')) + RotatedAxis() + xlab('Marker Genes')+ ylab('Clusters')

BCC.cluster.markers

## Dimensions: width = 1500px, height = 750px
## Saved as BCC_Cluster_Markers.jpeg

```



BCC Clustering Comparison Heatmap

```{r}

load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 123813\\BCCIOrawdata")

BCCIO.meta.data1 <- BCCIO.meta.data[order(BCCIO.meta.data[, "sort"], -xtfrm(BCCIO.meta.data[, "treatment"])),]
BCCIO.meta.data2 <- BCCIO@meta.data[order(BCCIO@meta.data[, "sort"]),]

BCCIO.meta.data1["rownames"] <- rownames(BCCIO.meta.data1)
BCCIO.meta.data2["rownames"] <- rownames(BCCIO.meta.data2)

BCCIO.meta.data.merge <- merge(BCCIO.meta.data1, BCCIO.meta.data2, by.x=c("rownames"), by.y=c("rownames"))
BCCIO.clustering.comparison <- BCCIO.meta.data.merge[c("cluster", "seurat_clusters")]

uniqueclusters1 <- as.matrix(unique(BCCIO.clustering.comparison[,1]))
uniqueclusters2 <- as.matrix(unique(BCCIO.clustering.comparison[,2]))

BCCIO.distribution.table <- dist_table(BCCIO.clustering.comparison, uniqueclusters1, uniqueclusters2, 1, 2)

BCCIO.proportion.table <- prop_table(BCCIO.distribution.table, uniqueclusters1, uniqueclusters2)

clusterlabels1 <- c('CD8+ memory T cells', 'CD8+ exhausted T cells', 'Regulatory T cells', 'CD4+ T cells', 'Proliferating T cells', 'NK cells', 'B cells 1', 'CD8+ activated T cells', 'Macrophages', 'pDCs', 'Plasma cells', 'Dendritic cells', 'CAFs', 'Endothelial cells', 'Myofibroblasts', 'B cells 2', 'Melanocytes', 'Malignant cells 1', 'Malignant cells 2')

clusterlabels2 <- c('CD8+ memory T cells', 'CD8+ exhausted T cells', 'Regulatory T cells', 'CD4+ T cells', 'Proliferating T cells', 'NK cells', 'Other', 'CD8+ activated T cells', 'B cells', 'Macrophages', 'pDCs', 'Plasma cells', 'Dendritic cells', 'CAFs', 'Endothelial cells', 'Myofibroblasts', 'Melanocytes', 'Malignant cells 1', 'Malignant cells 2')

BCCIO.heatmap <- ggplot_heatmap(BCCIO.proportion.table, clusterlabels1, clusterlabels2, c("Metadata Clustering"), c("Unsupervised Clustering")) + 
  theme(axis.text = element_text(size=10), axis.title = element_text(size=14)) + 
  scale_fill_viridis(discrete=FALSE, name = "Proportion of \nsimilar cells")

BCCIO.heatmap

## Dimensions: width = 1000px (scale up)
## Saved as BCC_Clustering_Heatmap.jpeg

```



BCC Pretreatment vs. Posttreatment Barcharts 

```{r}

## Separate into pretreatment and posttreatment batches

Idents(BCCIO) <- BCCIO$treatment
BCCIO.pre <- subset(BCCIO, idents=c("pre"))
BCCIO.post <- subset(BCCIO, idents=c("post"))

## Pre-treatment barchart

dataframe <- data.frame("Batch" = BCCIO.pre@meta.data$patient, "Clusters" = BCCIO.pre@meta.data$seurat_clusters)
tibble <- as_tibble(dataframe)
  
summary <- tibble %>% group_by(Clusters, Batch) %>% count()

summary$cluster <- as.vector(summary$Clusters)

summary[summary == "CD8_memory"] <- "CD8+ memory T cells"
summary[summary == "CD4"] <- "CD4+ T cells"
summary[summary == "Tregs"] <- "Regulatory T cells"
summary[summary == "CD8_activation"] <- "CD8+ activated T cells"
summary[summary == "CD8_exhausted"] <- "CD8+ exhausted T cells"
summary[summary == "Plasma"] <- "Plasma cells"
summary[summary == "Tumor 1"] <- "Malignant cells 1"
summary[summary == "Tumor 2"] <- "Malignant cells 2"
summary[summary == "T_proliferation"] <- "Proliferating T cells"
summary[summary == "NK"] <- "NK cells"
summary[summary == "Endothelial"] <- "Endothelial cells"
summary[summary == "DCs"] <- "Dendritic cells"
summary[summary == "Misc."] <- "Other"

summary$Clusters <- as.factor(summary$cluster)
summary <- subset(summary, select = -cluster)

normalized_dataset <- summary %>% group_by(Batch) %>% mutate(normalize = n/sum(n))

normalized_dataset$Clusters <- factor(normalized_dataset$Clusters,
                                      levels = c("CD4+ T cells", "CD8+ memory T cells", "CD8+ activated T cells", "CD8+ exhausted T cells", "Regulatory T cells", "Proliferating T cells", "NK cells", "B cells", "Macrophages", "Dendritic cells", "pDCs", "Plasma cells", "CAFs", "Endothelial cells", "Myofibroblasts", "Melanocytes", "Malignant cells 1", "Malignant cells 2", "Other"))

BCC.pre.patient.barchart <- ggplot(normalized_dataset, aes(x=Clusters, y=normalize, fill=Batch)) + 
      geom_col(position = 'fill') + 
      scale_fill_viridis(discrete=TRUE) +
      theme_classic() +
      theme(axis.text.x = element_text(angle=45, hjust=1)) + 
      ylab("Cell Proportion Normalized by Batch") + 
      ggtitle("Pre-treatment") +
      theme(plot.title = element_text(hjust = 0.5))

## Post-treatment barchart

dataframe <- data.frame("Batch" = BCCIO.post@meta.data$patient, "Clusters" = BCCIO.post@meta.data$seurat_clusters)
tibble <- as_tibble(dataframe)
  
summary <- tibble %>% group_by(Clusters, Batch) %>% count()

summary$cluster <- as.vector(summary$Clusters)

summary[summary == "CD8_memory"] <- "CD8+ memory T cells"
summary[summary == "CD4"] <- "CD4+ T cells"
summary[summary == "Tregs"] <- "Regulatory T cells"
summary[summary == "CD8_activation"] <- "CD8+ activated T cells"
summary[summary == "CD8_exhausted"] <- "CD8+ exhausted T cells"
summary[summary == "Plasma"] <- "Plasma cells"
summary[summary == "Tumor 1"] <- "Malignant cells 1"
summary[summary == "Tumor 2"] <- "Malignant cells 2"
summary[summary == "T_proliferation"] <- "Proliferating T cells"
summary[summary == "NK"] <- "NK cells"
summary[summary == "Endothelial"] <- "Endothelial cells"
summary[summary == "DCs"] <- "Dendritic cells"
summary[summary == "Misc."] <- "Other"

summary$Clusters <- as.factor(summary$cluster)
summary <- subset(summary, select = -cluster)

normalized_dataset <- summary %>% group_by(Batch) %>% mutate(normalize = n/sum(n))

normalized_dataset$Clusters <- factor(normalized_dataset$Clusters,
                                      levels = c("CD4+ T cells", "CD8+ memory T cells", "CD8+ activated T cells", "CD8+ exhausted T cells", "Regulatory T cells", "Proliferating T cells", "NK cells", "B cells", "Macrophages", "Dendritic cells", "pDCs", "Plasma cells", "CAFs", "Endothelial cells", "Myofibroblasts", "Melanocytes", "Malignant cells 1", "Malignant cells 2", "Other"))

BCC.post.patient.barchart <- ggplot(normalized_dataset, aes(x=Clusters, y=normalize, fill=Batch)) + 
      geom_col(position = 'fill') + 
      theme_classic() +
      theme(axis.text.x = element_text(angle=45, hjust=1)) + 
      ylab("Cell Proportion Normalized by Patient") + 
      scale_fill_viridis(discrete=TRUE, name = "Patient") + 
      ggtitle("Post-treatment") +
      theme(plot.title = element_text(hjust = 0.5))


BCC.pre.patient.barchart + NoLegend() +
  BCC.post.patient.barchart

## Dimensions: width= 1000px, height = 500px
## Saved as BCC_Barcharts_Treatment.jpeg

```



BCC Responder vs. Non-responder Barcharts

```{r}

## Separate into pretreatment and posttreatment batches

Idents(BCCIO) <- BCCIO$patient
BCCIO.responders <- subset(BCCIO, idents=c("su001", "su002", "su003", "su004", "su009", "su012"))
BCCIO.nonresponders <- subset(BCCIO, idents=c("su005", "su006", "su007", "su008", "su010"))

## Responders barchart

dataframe <- data.frame("Batch" = BCCIO.responders@meta.data$treatment, "Clusters" = BCCIO.responders@meta.data$seurat_clusters)
tibble <- as_tibble(dataframe)
  
summary <- tibble %>% group_by(Clusters, Batch) %>% count()

summary$cluster <- as.vector(summary$Clusters)
summary$batches <- as.vector(summary$Batch)

summary[summary == "CD8_memory"] <- "CD8+ memory T cells"
summary[summary == "CD4"] <- "CD4+ T cells"
summary[summary == "Tregs"] <- "Regulatory T cells"
summary[summary == "CD8_activation"] <- "CD8+ activated T cells"
summary[summary == "CD8_exhausted"] <- "CD8+ exhausted T cells"
summary[summary == "Plasma"] <- "Plasma cells"
summary[summary == "Tumor 1"] <- "Malignant cells 1"
summary[summary == "Tumor 2"] <- "Malignant cells 2"
summary[summary == "T_proliferation"] <- "Proliferating T cells"
summary[summary == "NK"] <- "NK cells"
summary[summary == "Endothelial"] <- "Endothelial cells"
summary[summary == "DCs"] <- "Dendritic cells"
summary[summary == "Misc."] <- "Other"
summary[summary == "post"] <- "Post-treatment"
summary[summary == "pre"] <- "Pre-treatment"

summary$Clusters <- as.factor(summary$cluster)
summary <- subset(summary, select = -cluster)
summary$Batch <- as.factor(summary$batches)
summary <- subset(summary, select = -batches)

normalized_dataset <- summary %>% group_by(Batch) %>% mutate(normalize = n/sum(n))

normalized_dataset$Clusters <- factor(normalized_dataset$Clusters,
                                      levels = c("CD4+ T cells", "CD8+ memory T cells", "CD8+ activated T cells", "CD8+ exhausted T cells", "Regulatory T cells", "Proliferating T cells", "NK cells", "B cells", "Macrophages", "Dendritic cells", "pDCs", "Plasma cells", "CAFs", "Endothelial cells", "Myofibroblasts", "Melanocytes", "Malignant cells 1", "Malignant cells 2", "Other"))

normalized_dataset$Batch <- factor(normalized_dataset$Batch, 
                                   levels = c("Pre-treatment", "Post-treatment"))

BCC.responders.barchart <- ggplot(normalized_dataset, aes(x=Clusters, y=normalize, fill=Batch)) + 
      geom_col(position = 'fill') + 
      scale_fill_manual(values = c("royalblue3", "lightskyblue1")) +
      theme_classic() +
      theme(axis.text.x = element_text(angle=45, hjust=1)) + 
      ylab("Cell Proportion Normalized by Batch") + 
      ggtitle("Responders") +
      theme(plot.title = element_text(hjust = 0.5))

## Nonresponders barchart

dataframe <- data.frame("Batch" = BCCIO.nonresponders@meta.data$treatment, "Clusters" = BCCIO.nonresponders@meta.data$seurat_clusters)
tibble <- as_tibble(dataframe)
  
summary <- tibble %>% group_by(Clusters, Batch) %>% count()

summary$cluster <- as.vector(summary$Clusters)
summary$batches <- as.vector(summary$Batch)

summary[summary == "CD8_memory"] <- "CD8+ memory T cells"
summary[summary == "CD4"] <- "CD4+ T cells"
summary[summary == "Tregs"] <- "Regulatory T cells"
summary[summary == "CD8_activation"] <- "CD8+ activated T cells"
summary[summary == "CD8_exhausted"] <- "CD8+ exhausted T cells"
summary[summary == "Plasma"] <- "Plasma cells"
summary[summary == "Tumor 1"] <- "Malignant cells 1"
summary[summary == "Tumor 2"] <- "Malignant cells 2"
summary[summary == "T_proliferation"] <- "Proliferating T cells"
summary[summary == "NK"] <- "NK cells"
summary[summary == "Endothelial"] <- "Endothelial cells"
summary[summary == "DCs"] <- "Dendritic cells"
summary[summary == "Misc."] <- "Other"
summary[summary == "post"] <- "Post-treatment"
summary[summary == "pre"] <- "Pre-treatment"

summary$Clusters <- as.factor(summary$cluster)
summary <- subset(summary, select = -cluster)
summary$Batch <- as.factor(summary$batches)
summary <- subset(summary, select = -batches)

normalized_dataset <- summary %>% group_by(Batch) %>% mutate(normalize = n/sum(n))

normalized_dataset$Clusters <- factor(normalized_dataset$Clusters,
                                      levels = c("CD4+ T cells", "CD8+ memory T cells", "CD8+ activated T cells", "CD8+ exhausted T cells", "Regulatory T cells", "Proliferating T cells", "NK cells", "B cells", "Macrophages", "Dendritic cells", "pDCs", "Plasma cells", "CAFs", "Endothelial cells", "Myofibroblasts", "Melanocytes", "Malignant cells 1", "Malignant cells 2", "Other"))

normalized_dataset$Batch <- factor(normalized_dataset$Batch, 
                                   levels = c("Pre-treatment", "Post-treatment"))

BCC.nonresponders.barchart <- ggplot(normalized_dataset, aes(x=Clusters, y=normalize, fill=Batch)) + 
      geom_col(position = 'fill') + 
      scale_fill_manual(values = c("tomato4", "salmon1")) +
      theme_classic() +
      theme(axis.text.x = element_text(angle=45, hjust=1)) + 
      ylab("Cell Proportion Normalized by Batch") + 
      ggtitle("Nonresponders") +
      theme(plot.title = element_text(hjust = 0.5))


BCC.responders.barchart + BCC.nonresponders.barchart

## Dimensions: width= 1200px, height = 500px
## Post-processing to modify legend in Powerpoint
## Saved as BCC_Barcharts_Responders.jpeg

```



BCC Macrophage Pretreatment Enhanced Volcano Plot

```{r}





```
