---
title: "PDAC Analysis"
output: html_notebook
---

```{r}
## Load packages
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(pracma)
library(RColorBrewer)
library(EnhancedVolcano)

## Load functions
source("ClusteringAnalysisFunctions.R", echo=FALSE)
source("SeuratClusteringFunctions.R", echo=FALSE)
source("DifferentialExpressionFunctions.R", echo=FALSE)

## Load PDAC integrated object
load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC155698.integrated")

```



Analyze GSE 111672 Datasets

```{r}

pdac1.read <- read.table("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\GSE111672_PDAC-A-indrop-filtered-expMat.txt", sep = "\t", as.is = 1)
pdac1.read <- pdac1.read[-c(9946, 9947), ] ## Duplicates of 1-Mar and 2-Mar genes (excel error??)
rownames(pdac1.read) <- pdac1.read$V1 ## Rownames are stored in the first column of the dataset
pdac1.read <- pdac1.read[, -c(1)] ## Delete extra rownames vector
names(pdac1.read) <- lapply(pdac1.read[1,], as.character) ## Column names are stored in the first row of the dataset
pdac1.read <- pdac1.read[-1,] ## Delete extra colnames vector

pdac2.read <- read.table("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\GSE111672_PDAC-B-indrop-filtered-expMat.txt", sep = "\t", header = TRUE, as.is = 1)
pdac2.read <- pdac2.read[-c(9946, 9947), ]
rownames(pdac2.read) <- pdac2.read$Genes
pdac2.read <- pdac2.read[, -c(1)]

pdac1 <- CreateSeuratObject(counts=pdac1.read, project="PDAC1", min.cells=1, min.features=1)
pdac2 <- CreateSeuratObject(counts=pdac2.read, project="PDAC2", min.cells=1, min.features=1)
```



Initial PDAC Datasets Clustering

```{r}

pdac1 <- preprocessing(pdac1, heatmap=FALSE)
pdac2 <- preprocessing(pdac2, heatmap=FALSE)

pdac1 <- clustering(pdac1, PCs=50, resolution=0.5)
pdac2 <- clustering(pdac2, PCs=50, resolution=0.5)


pdac1.umap <- DimPlot(pdac1, reduction="umap", label=TRUE, pt.size=0.3) + NoLegend()
pdac2.umap <- DimPlot(pdac2, reduction="umap", label=TRUE, pt.size=0.3) + NoLegend()
pdac1.umap + pdac2.umap

## Checkpoint

save(pdac1, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\pdac1")
save(pdac2, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\pdac2")

load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\pdac1")
load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\pdac2")
```



PDAC Datasets Clustering Comparison 

```{r}

pdac1.read.cell.labels <- read.table("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\GSE111672_PDAC-A-indrop-filtered-expMat.txt", nrows=1, header=FALSE, sep = "\t", as.is = 1)
pdac1.read.cell.labels <- t(pdac1.read.cell.labels)
pdac1.read.cell.labels <- pdac1.read.cell.labels[-c(1),]
pdac1@meta.data$orig_clusters <- pdac1.read.cell.labels

pdac2.read.cell.labels <- read.table("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\GSE111672_PDAC-B-indrop-filtered-expMat.txt", nrows=1, header=FALSE, sep = "\t", as.is = 1)
pdac2.read.cell.labels <- t(pdac2.read.cell.labels)
pdac2.read.cell.labels <- pdac2.read.cell.labels[-c(1),]
pdac2@meta.data$orig_clusters <- pdac2.read.cell.labels


pdac1.umap.orig <- DimPlot(pdac1, group.by = "orig_clusters", label=TRUE, pt.size=0.2) + NoLegend()
pdac2.umap.orig <- DimPlot(pdac2, group.by = "orig_clusters", label=TRUE, pt.size=0.2) + NoLegend()

pdac1.umap.orig
pdac2.umap.orig
pdac1.idents <- c("Ductal 1", "Ductal 2", "Ductal 3", "Cancer clone B", "Macrophages", "Cancer clone A", "Ductal 4", "T & NK", "Tuft", "Endocrine & RBCs", "Acinar", "Endothelial")

pdac2.idents <- c("Ductal 1", "Ductal 2", "Cancer clone A", "Ductal 3", "Endothelial", "Macrophages", "Ductal 4", "Tuft & Endocrine", "Acinar")

names(pdac1.idents) <- levels(pdac1)
pdac1 <- RenameIdents(pdac1, pdac1.idents)

names(pdac2.idents) <- levels(pdac2)
pdac2 <- RenameIdents(pdac2, pdac2.idents)

```



```{r}

pdac1 <- FindNeighbors(pdac1, dims = 1:50)
pdac1 <- FindClusters(pdac1, resolution = 0.5, algorithm = 3)
pdac1.markers <- FindAllMarkers(pdac1, test.use = 'wilcox', min.pct = 0.25, logfc.threshold = 0.50, max.cells.per.ident = 500)
pdac1.top25 <- pdac1.markers %>% group_by(cluster) %>% top_n(n=25, wt=avg_logFC)

pdac2 <- FindNeighbors(pdac2, dims = 1:50)
pdac2 <- FindClusters(pdac2, resolution = 0.5, algorithm = 3)
pdac2.markers <- FindAllMarkers(pdac2, test.use = 'wilcox', min.pct = 0.25, logfc.threshold = 0.50, max.cells.per.ident = 500)
pdac2.top25 <- pdac2.markers %>% group_by(cluster) %>% top_n(n=25, wt=avg_logFC)

```



Combined analysis of pdac datasets

```{r}
## Merge and process datasets 
pdac.combined <- merge(pdac1, y=pdac2, add.cell.ids=c("Patient 1", "Patient 2"), project="Merged PDAC")
pdac.combined <- preprocessing(pdac.combined)
pdac.combined <- clustering(pdac.combined)

## Create UMAPS
pdac.combined.umap <- DimPlot(pdac.combined, reduction="umap")
pdac.combined.identsep.umap <- DimPlot(pdac.combined, reduction="umap", group.by="orig.ident")
pdac.combined.umap + pdac.combined.identsep.umap


```



PDAC Merged Heatmap Comparison 

```{r}
pdac.combined.meta.data <- pdac.combined@meta.data
pdac1.meta.data <- pdac1@meta.data
pdac2.meta.data <- pdac2@meta.data

pdac1.meta.data$rownames <- rownames(pdac1.meta.data)
pdac1.meta.data$rownames <- paste0("Patient 1_", pdac1.meta.data$rownames)

pdac2.meta.data$rownames <- rownames(pdac2.meta.data)
pdac2.meta.data$rownames <- paste0("Patient 2_", pdac2.meta.data$rownames)
pdac2.meta.data$seurat_clusters <- paste0("B", pdac2.meta.data$seurat_clusters)

pdac.combined.meta.data$rownames <- rownames(pdac.combined.meta.data)
mergetest1 <- merge(pdac.combined.meta.data, pdac1.meta.data, by=c("rownames", "nCount_RNA"))
mergetest2 <- merge(pdac.combined.meta.data, pdac2.meta.data, by=c("rownames", "nCount_RNA"))
pdac.combined.merged <- rbind(mergetest1, mergetest2)

uniqueclusters1 <- as.matrix(unique(pdac.combined.merged$seurat_clusters.x))
uniqueclusters2 <- as.matrix(unique(pdac.combined.merged$seurat_clusters.y))

pdac.distribution.table <- dist_table(pdac.combined.merged, uniqueclusters1, uniqueclusters2, 6, 12)

pdac.proportions.table <- prop_table(pdac.distribution.table, uniqueclusters1, uniqueclusters2)

uniqueclusters2 <- c("Acinar", "Cancer clone A", "Ductal 4", "Cancer clone B", "Ductal 3", "Ductal 1", "Ductal 2", "Tuft", "T & NK", "Endothelial", "Macrophages", "Endocrine & RBCs", "B_Acinar", "B_Cancer clone A", "B_Ductal 4", "B_Ductal 3", "B_Ductal 2", "B_Tuft & Endocrine", "B_Macrophages", "B_Ductal 1", "B_Endothelial")

pdac.comparison.heatmap <- ggplot_heatmap(pdac.proportions.table, uniqueclusters1, uniqueclusters2, c("RdGy"), c("Merged Clustering"), c("Individual Clustering"))
  
pdac.comparison.heatmap

```



Integrated analysis of datasets 

```{r}

pdac.list <- list(pdac1, pdac2)
pdac.anchors <- FindIntegrationAnchors(object.list=pdac.list, dims=1:50)
pdac.integrated <- IntegrateData(anchorset=pdac.anchors, dims=1:50)
DefaultAssay(pdac.integrated) <- "RNA"

pdac.integrated <- preprocessing(pdac.integrated, heatmap=FALSE)
pdac.integrated <- clustering(pdac.integrated, PCs=50, genes=FALSE)
pdac.integrated <- RunUMAP(pdac.integrated, dims=1:50)
pdac.integrated.heatmap <- DimPlot(pdac.integrated, reduction="umap", label=TRUE, pt.size = 0.1) + NoLegend()

labels <- c("Ductal 1", "Ductal 2", "Ductal 3", "Cancer clone A (pat. B)", "Ductal 4", "Cancer clone B (pat. A)", "Endothelial", "Macrophages", "Cancer clone A (pat. A)", "NK & T 1", "Tuft", "NK & T 2", "Acinar", "Endocrine & RBCs")

names(labels) <- levels(pdac.integrated)
pdac.integrated <- RenameIdents(pdac.integrated, labels)
pdac.integrated@meta.data$seurat_clusters <- pdac.integrated@active.ident
DimPlot(pdac.integrated, reduction="umap", label=TRUE, pt.size = 0.1) + NoLegend() 


##Checkpoint
save(pdac.integrated, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\pdac.integrated")
load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 111672\\pdac.integrated")


p1 <- DimPlot(pdac.integrated, reduction="umap", group.by="orig.ident")
p2 <- DimPlot(pdac.integrated, reduction="umap", group.by="seurat_clusters")
p3 <- DimPlot(pdac.integrated, reduction="umap", group.by="orig_clusters", label=TRUE) + NoLegend()
p1 + p3

```



PDAC Integrated Heatmap Analysis

```{r}

pdac.integrated.meta.data <- pdac.integrated@meta.data
pdac1.meta.data <- pdac1@meta.data
pdac2.meta.data <- pdac2@meta.data

pdac1.meta.data$rownames <- rownames(pdac1.meta.data)
pdac1.meta.data$rownames <- paste0("Patient 1_", pdac1.meta.data$rownames)

pdac2.meta.data$rownames <- rownames(pdac2.meta.data)
pdac2.meta.data$rownames <- paste0("Patient 2_", pdac2.meta.data$rownames)
pdac2.meta.data$seurat_clusters <- as.factor(as.numeric(as.character(pdac2.meta.data$seurat_clusters)) + 12)

pdac.integrated.meta.data$rownames <- rownames(pdac.combined.meta.data)
integratetest1 <- merge(pdac.integrated.meta.data, pdac1.meta.data, by=c("rownames", "nCount_RNA"))
integratetest2 <- merge(pdac.integrated.meta.data, pdac2.meta.data, by=c("rownames", "nCount_RNA"))
pdac.integrated.merged <- rbind(integratetest1, integratetest2)

uniqueclusters1 <- as.matrix(unique(pdac.integrated.merged$seurat_clusters.x))
uniqueclusters2 <- as.matrix(unique(pdac.integrated.merged$seurat_clusters.y))

pdac.integrated.distribution.table <- dist_table(pdac.integrated.merged, uniqueclusters1, uniqueclusters2, 6, 12)

pdac.integrated.proportions.table <- prop_table(pdac.integrated.distribution.table, uniqueclusters1, uniqueclusters2)

uniqueclusters2 <- c("Acinar", "Cancer clone A", "Ductal 4", "Cancer clone B", "Ductal 3", "Ductal 1", "Ductal 2", "Tuft", "T & NK", "Endothelial", "Macrophages", "Endocrine & RBCs", "B_Acinar", "B_Cancer clone A", "B_Ductal 4", "B_Ductal 3", "B_Ductal 2", "B_Tuft & Endocrine", "B_Macrophages", "B_Ductal 1", "B_Endothelial")

pdac.integrated.comparison.heatmap <- ggplot_heatmap(pdac.integrated.proportions.table, uniqueclusters1, uniqueclusters2, xlab=c("Integrated Clustering"), ylab=c("Individual Clustering"))

pdac.integrated.comparison.heatmap

```



PDAC Gene Expression Analysis 

```{r}

Idents(pdac.integrated) <- pdac.integrated$seurat_clusters

pdac.integrated.vlnplot.PDCD1 <- VlnPlot(pdac.integrated, "PDCD1") + NoLegend()
pdac.integrated.vlnplot.CD274 <- VlnPlot(pdac.integrated, "CD274") + NoLegend()
pdac.integrated.vlnplot.PDCD1LG2 <- VlnPlot(pdac.integrated, "PDCD1LG2") + NoLegend()
pdac.integrated.vlnplot.TCF7 <- VlnPlot(pdac.integrated, "TCF7") + NoLegend()
pdac.integrated.vlnplot.PDCD1 + pdac.integrated.vlnplot.CD274 + pdac.integrated.vlnplot.PDCD1LG2 + pdac.integrated.vlnplot.TCF7



```



```{r}

pdac.integrated.treatment.barchart <- ggplot_barcharts(pdac.integrated@meta.data$orig.ident, pdac.integrated@meta.data$seurat_clusters, colors="Accent")
pdac.integrated.treatment.barchart


```


```{r}

set.seed(001)
pdac.subset <- SubsetData(pdac.integrated, max.cells.per.ident=100)
Idents(pdac.subset) = pdac.subset$orig.ident
pdac.averages <- AverageExpression(pdac.subset)
head(pdac.averages[["RNA"]][c("TCF7", "PDCD1", "CD274", "PDCD1LG2"),])

```




```{r}
features = c("IL7R", "CCR7", "S100A4", "CD14", "LYZ", "MS4A1", "CD8A", "FCGR3A", "MS4A7", "GNLY", "NKG7", "FCER1A", "CST3", "PPBP")
FeaturePlot(pdac.combined, features=features)
DotPlot(pdac.combined, features=features) + RotatedAxis()
RidgePlot(pdac.combined, features=features, ncol=6)
DoHeatmap(subset(pdac.combined, downsample=100), features=features)
```



##############################################################################################

Analyze GSE 155698 Dataset

```{r}

PDAC155698.1.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710689_PDAC_TISSUE_1\\PDAC_TISSUE_1\\filtered_feature_bc_matrix\\")
PDAC155698.2.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710690_PDAC_TISSUE_2\\PDAC_TISSUE_2\\filtered_feature_bc_matrix\\")
PDAC155698.3.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710691_PDAC_TISSUE_3\\PDAC_TISSUE_3\\filtered_feature_bc_matrix\\")
PDAC155698.4.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710692_PDAC_TISSUE_4\\PDAC_TISSUE_4\\filtered_feature_bc_matrix\\")
PDAC155698.5.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710693_PDAC_TISSUE_5\\PDAC_TISSUE_5\\filtered_feature_bc_matrix\\")
PDAC155698.6.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710694_PDAC_TISSUE_6\\PDAC_TISSUE_6\\filtered_feature_bc_matrix\\")
PDAC155698.7.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710695_PDAC_TISSUE_7\\PDAC_TISSUE_7\\filtered_feature_bc_matrix\\")
PDAC155698.8.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710696_PDAC_TISSUE_8\\PDAC_TISSUE_8\\filtered_feature_bc_matrix\\")
PDAC155698.9.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710697_PDAC_TISSUE_9\\PDAC_TISSUE_9\\filtered_feature_bc_matrix\\")
PDAC155698.10.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710698_PDAC_TISSUE_10\\PDAC_TISSUE_10\\filtered_feature_bc_matrix\\")
PDAC155698.11A.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710699_PDAC_TISSUE_11A\\PDAC_TISSUE_11A\\filtered_feature_bc_matrix\\")
PDAC155698.11B.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710700_PDAC_TISSUE_11B\\PDAC_TISSUE_11B\\filtered_feature_bc_matrix\\")
PDAC155698.12.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710701_PDAC_TISSUE_12\\PDAC_TISSUE_12\\filtered_feature_bc_matrix\\")
PDAC155698.13.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710702_PDAC_TISSUE_13\\PDAC_TISSUE_13\\filtered_feature_bc_matrix\\")
## PDAC155698.14.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710703_PDAC_TISSUE_14\\PDAC_TISSUE_14\\filtered_feature_bc_matrix\\")
PDAC155698.15.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710704_PDAC_TISSUE_15\\PDAC_TISSUE_15\\filtered_feature_bc_matrix\\")
PDAC155698.16.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC Samples\\GSM4710705_PDAC_TISSUE_16\\PDAC_TISSUE_16\\filtered_gene_bc_matrices\\")

## PDAC 14 file could not be opened 

PDAC155698.1 <- CreateSeuratObject(PDAC155698.1.read, project="PDAC155698.1", min.cells=3, min.features=200)
PDAC155698.2 <- CreateSeuratObject(PDAC155698.2.read, project="PDAC155698.2", min.cells=3, min.features=200)
PDAC155698.3 <- CreateSeuratObject(PDAC155698.3.read, project="PDAC155698.3", min.cells=3, min.features=200)
PDAC155698.4 <- CreateSeuratObject(PDAC155698.4.read, project="PDAC155698.4", min.cells=3, min.features=200)
PDAC155698.5 <- CreateSeuratObject(PDAC155698.5.read, project="PDAC155698.5", min.cells=3, min.features=200)
PDAC155698.6 <- CreateSeuratObject(PDAC155698.6.read, project="PDAC155698.6", min.cells=3, min.features=200)
PDAC155698.7 <- CreateSeuratObject(PDAC155698.7.read, project="PDAC155698.7", min.cells=3, min.features=200)
PDAC155698.8 <- CreateSeuratObject(PDAC155698.8.read, project="PDAC155698.8", min.cells=3, min.features=200)
PDAC155698.9 <- CreateSeuratObject(PDAC155698.9.read, project="PDAC155698.9", min.cells=3, min.features=200)
PDAC155698.10 <- CreateSeuratObject(PDAC155698.10.read, project="PDAC155698.10", min.cells=3, min.features=200)
PDAC155698.11A <- CreateSeuratObject(PDAC155698.11A.read, project="PDAC155698.11A", min.cells=3, min.features=200)
PDAC155698.11B <- CreateSeuratObject(PDAC155698.11B.read, project="PDAC155698.11B", min.cells=3, min.features=200)
PDAC155698.12 <- CreateSeuratObject(PDAC155698.12.read, project="PDAC155698.12", min.cells=3, min.features=200)
PDAC155698.13 <- CreateSeuratObject(PDAC155698.13.read, project="PDAC155698.13", min.cells=3, min.features=200)
## PDAC155698.14 <- CreateSeuratObject(PDAC155698.14.read, project="PDAC155698.14", min.cells=3, min.features=200)
PDAC155698.15 <- CreateSeuratObject(PDAC155698.15.read, project="PDAC155698.15", min.cells=3, min.features=200)
PDAC155698.16 <- CreateSeuratObject(PDAC155698.16.read, project="PDAC155698.16", min.cells=3, min.features=200)

## Merge datasets together

PDAC.merged <- merge(PDAC155698.1, y=c(PDAC155698.2, PDAC155698.3, PDAC155698.4, PDAC155698.5, PDAC155698.6, PDAC155698.7, PDAC155698.8, PDAC155698.9, PDAC155698.10, PDAC155698.11A, PDAC155698.11B, PDAC155698.12, PDAC155698.13, PDAC155698.15, PDAC155698.16), add.cell.ids = c("tissue1", "tissue2", "tissue3", "tissue4", "tissue5", "tissue6", "tissue7", "tissue8", "tissue9", "tissue10", "tissue11A", "tissue11B", "tissue12", "tissue13", "tissue15", "tissue16"), project="PDAC155698")


## Integrate datasets together

PDAC155698.list <- list(PDAC155698.1, PDAC155698.2, PDAC155698.3, PDAC155698.4, PDAC155698.5, PDAC155698.6, PDAC155698.7, PDAC155698.8, PDAC155698.9, PDAC155698.10, PDAC155698.11A, PDAC155698.11B, PDAC155698.12, PDAC155698.13, PDAC155698.15, PDAC155698.16)
PDAC155698.anchors <- FindIntegrationAnchors(object.list=PDAC155698.list, dims=1:50)
PDAC155698.integrated <- IntegrateData(anchorset=PDAC 155698.anchors, dims=1:50)
DefaultAssay(PDAC155698.integrated) <- "RNA"

```



Initial Analysis of GSE155698

```{r}
PDAC155698.integrated <- preprocessing(PDAC155698.integrated)

## Determine optimal resolution 
test_resolution(PDAC155698.integrated, 0.3)
test_resolution(PDAC155698.integrated, 0.4)
test_resolution(PDAC155698.integrated, 0.2)
test_resolution(PDAC155698.integrated, 0.5)
## Best resolution: 0.5 (27 clusters)

PDAC155698.integrated <- clustering(PDAC155698.integrated, resolution=0.5, genes=FALSE)

save(PDAC155698.integrated, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC155698.integrated")
memory.limit(size=160000000)

load("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC155698.integrated")

PDAC155698.integrated.umap <- DimPlot(PDAC155698.integrated, reduction="umap", label=TRUE)
PDAC155698.integrated.umap

```



Clustering Labels

```{r}

## Features taken from Figure 4B

FeaturePlot(PDAC155698.integrated, features=c('LAMP3', 'CCL22', 'TFF1', 'KRT18', 'KRT19', 'KRT8', 'CLU', 'MMP7', 'SPP1', 'REG1A', 'CTRB2', 'PRSS1', 'DCN', 'LUM', 'CPA3', 'TPSAB1', 'CDH5', 'VWF', 'PLVAP', 'IRF7', 'RGS5', 'PDGFRB', 'CD3E', 'NCAM1', 'NKG7', 'CD3D', 'CD14', 'HLA-DRA', 'GZMB', 'ITGAX', 'ITGAM', 'APOE', 'LYZ', 'IGJ', 'CD79A', 'MS4A1'))

labels <- c("0", "1", "Epithelial + Endocrine", "Macrophages", "Granulocytes", "Epithelial + Endocrine", "Dendritic", "Macrophages", "Epithelial + Endocrine", "Epithelial + Endocrine", "Fibroblasts", "Mast cells", "Macrophages", "Plasma cells", "Epithelial + Endocrine", "B cells", "Granulocytes", "Pericytes", "Acinar cells", "Epithelial + Endocrine", "20", "Acinar cells", "Epithelial + Endocrine", "Endothelial", "Macrophages", "Epithelial + Endocrine", "Plasma cells")

names(labels) <- levels(PDAC155698.integrated)
PDAC155698.integrated <- RenameIdents(PDAC155698.integrated, labels)
PDAC155698.integrated@meta.data$seurat_clusters <- PDAC155698.integrated@active.ident
DimPlot(PDAC155698.integrated, reduction="umap", label=TRUE, pt.size = 0.1) + NoLegend() 

## Separate T and NK cell types

PDAC155698.integrated.subset1 <- subset(PDAC155698.integrated, idents=c('0', '1', '20'))

PDAC155698.integrated.subset1 <- clustering(PDAC155698.integrated.subset1, resolution=1)
PDAC155698.integrated.subset1.umap <- DimPlot(PDAC155698.integrated.subset1, reduction="umap", label=TRUE)
PDAC155698.integrated.subset1.umap

FeaturePlot(PDAC155698.integrated.subset1, features=c("CD3E", "NCAM1", "NKG7", "CD3D", "FOXP3", "GZMB"))

labels <- c("CD4+ T cells", "Tregs", "CD4+ T cells", "CD8+ T cells", "CD8+ T cells", "CD8+ T cells", "NK cells", "CD4+ T cells", "NK cells", "CD8+ T cells", "Other", "CD4+ T cells", "CD8+ T cells", "CD8+ T cells")
names(labels) <- levels(PDAC155698.integrated.subset1)
PDAC155698.integrated.subset1 <- RenameIdents(PDAC155698.integrated.subset1, labels)

PDAC155698.integrated$sub_cluster <- as.character(Idents(PDAC155698.integrated))
PDAC155698.integrated$sub_cluster[Cells(PDAC155698.integrated.subset1)] <- paste(Idents(PDAC155698.integrated.subset1))
DimPlot(PDAC155698.integrated, group.by="sub_cluster", label = TRUE)

PDAC155698.integrated$seurat_clusters <- PDAC155698.integrated$sub_cluster
Idents(PDAC155698.integrated) <- PDAC155698.integrated$seurat_clusters

PDAC155698.integrated <- RenameIdents(PDAC155698.integrated, 'Other' = 'Macrophages')
PDAC155698.integrated.umap <- DimPlot(PDAC155698.integrated, reduction="umap", label=TRUE)
PDAC155698.integrated.umap

## Checkpoint

save(PDAC155698.integrated, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC155698.integrated")

PDAC155698.integrated.umap <- DimPlot(PDAC155698.integrated, reduction="umap", label=TRUE) + NoLegend()
PDAC155698.integrated.umap

```



Add patient labels

```{r}

Idents(PDAC155698.integrated) <- PDAC155698.integrated$orig.ident
PDAC155698.integrated@meta.data$patient[is.na(PDAC155698.integrated@meta.data$patient)] <- PDAC155698.integrated$meta.data$orig.ident

index = c("PDAC155698.1", "PDAC155698.2", "PDAC155698.3", "PDAC155698.4", "PDAC155698.5", "PDAC155698.6", "PDAC155698.7", "PDAC155698.8", "PDAC155698.9", "PDAC155698.10", "PDAC155698.11A", "PDAC155698.11B", "PDAC155698.12", "PDAC155698.13", "PDAC155698.15", "PDAC155698.16")

values <- c("pdac1", "pdac2", "pdac3", "pdac4", "pdac5", "pdac6", "pdac7", "pdac8", "pdac9", "pdac10", "pdac11A", "pdac11B", "pdac12", "pdac13", "pdac15", "pdac16")

PDAC155698.integrated@meta.data$patient <- values[match(PDAC155698.integrated@meta.data$orig.ident, index)]

save(PDAC155698.integrated, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC155698.integrated")

```



Distribution of samples in integrated clustering

```{r}

PDAC155698.tissue.umap <- DimPlot(PDAC155698.integrated, reduction="umap", group.by="orig.ident")
PDAC155698.tissue.umap

PDAC155698.tissue.barchart <- ggplot_barcharts(PDAC155698.integrated@meta.data$orig.ident, PDAC155698.integrated@meta.data$seurat_clusters, colors="Accent")
PDAC155698.tissue.barchart

```



Gene expression of PD-1 pathway

```{r}

set.seed(001)
pdac.subset <- PDAC155698.integrated[, sample(colnames(PDAC155698.integrated), size = 5000)]

pdac.subset$pdac <- "yes"

Idents(pdac.subset) <- pdac.subset$orig.ident
pdac.patient.averages <- AverageExpression(pdac.subset)
pdac.patient.averages.pd1 <- pdac.patient.averages[["RNA"]][c("TCF7", "PDCD1", "CD274", "PDCD1LG2"),]
pdac.patient.averages.pd1 <- as.data.frame(pdac.patient.averages.pd1)
pdac.patient.averages.pd1$mean <- rowMeans(pdac.patient.averages.pd1)
View(pdac.patient.averages.pd1)

Idents(pdac.subset) <- pdac.subset$pdac
pdac.total.averages <- AverageExpression(pdac.subset)
head(pdac.total.averages[["RNA"]][c("TCF7", "PDCD1", "CD274", "PDCD1LG2"),])

## Violin plots

PDAC.vlnplot.PD1 <- VlnPlot(pdac.subset, "PDCD1", pt.size = 0.1) + NoLegend()
PDAC.vlnplot.PDL1  <- VlnPlot(pdac.subset, "CD274", pt.size = 0.1) + NoLegend()
PDAC.vlnplot.PDL2  <- VlnPlot(pdac.subset, "PDCD1LG2", pt.size = 0.1) + NoLegend()
PDAC.vlnplot.TCF7  <- VlnPlot(pdac.subset, "TCF7", pt.size = 0.1) + NoLegend()

grid.arrange(PDAC.vlnplot.PD1, PDAC.vlnplot.PDL1, PDAC.vlnplot.PDL2, PDAC.vlnplot.TCF7, ncol=4)

```



GSE155698 Healthy Pancreas Samples

```{r}

## Read datasets

PDAC155698.H1.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\Healthy Samples\\GSM4710726_Healthy_PBMC_1\\Healthy_PBMC_1\\filtered_feature_bc_matrix")
PDAC155698.H2.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\Healthy Samples\\GSM4710727_Healthy_PBMC_2\\Healthy_PBMC_2\\filtered_feature_bc_matrix")
PDAC155698.H3.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\Healthy Samples\\GSM4710728_Healthy_PBMC_3\\Healthy_PBMC_3\\filtered_feature_bc_matrix")
PDAC155698.H4.read <- Read10X("C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\Healthy Samples\\GSM4710729_Healthy_PBMC_4\\Healthy_PBMC_4\\filtered_feature_bc_matrix")

## Create Seurat objects

PDAC155698.H1 <- CreateSeuratObject(PDAC155698.H1.read, project="PDAC155698.H1", min.cells=3, min.features=200)
PDAC155698.H2 <- CreateSeuratObject(PDAC155698.H2.read, project="PDAC155698.H2", min.cells=3, min.features=200)
PDAC155698.H3 <- CreateSeuratObject(PDAC155698.H3.read, project="PDAC155698.H3", min.cells=3, min.features=200)
PDAC155698.H4 <- CreateSeuratObject(PDAC155698.H4.read, project="PDAC155698.H4", min.cells=3, min.features=200)

## Integrate datasets

PDAC155698.all.list <- list(PDAC155698.1, PDAC155698.2, PDAC155698.3, PDAC155698.4, PDAC155698.5, PDAC155698.6, PDAC155698.7, PDAC155698.8, PDAC155698.9, PDAC155698.10, PDAC155698.11A, PDAC155698.11B, PDAC155698.12, PDAC155698.13, PDAC155698.15, PDAC155698.16, PDAC155698.H1, PDAC155698.H2, PDAC155698.H3, PDAC155698.H4)
PDAC155698.all.anchors <- FindIntegrationAnchors(object.list=PDAC155698.all.list, dims=1:50)
PDAC155698.all <- IntegrateData(anchorset=PDAC155698.all.anchors, dims=1:50)
DefaultAssay(PDAC155698.all) <- "RNA"

## Preprocessing

PDAC155698.all <- preprocessing(PDAC155698.all)
PDAC155698.all <- clustering(PDAC155698.all)

```



####################################################################################################################

Redo PDAC clustering

```{r}

## Load PDAC155698.integrated dataset

PDAC.integrated <- PDAC155698.integrated
remove(PDAC155698.integrated)

PDAC.integrated@meta.data$seurat_clusters <- PDAC.integrated@meta.data$RNA_snn_res.0.5

Idents(PDAC.integrated) <- PDAC.integrated$seurat_clusters
PDAC.integrated.umap <- DimPlot(PDAC.integrated, reduction="umap", label=TRUE)
PDAC.integrated.umap

```



Re-determine cluster labels for PDAC

```{r}

## Use features from the old PDAC paper

features <- c('CD79A', 'MS4A1', 'KRT18', 'KRT8', 'TFF1', 'LYZ', 'APOE', 'HLA-DRA', 'C1QA', 'CD14', 'CD3D', 'CD2', 'NKG7', 'IGFBP7', 'ACTA2', 'RGS5', 'PRSS1', 'CTRB2', 'REG1A', 'COL1A1', 'COL3A1', 'LUM', 'DCN', 'MMP7', 'CLU', 'SPP1', 'IGJ', 'IGLL5', 'IRF7', 'GZMB', 'CCL22', 'LAMP3', 'PLVAP', 'VWF', 'TPSAB1', 'CPA3', 'HBB', 'HBA2', 'KRT19', 'SPINK1')

DotPlot(PDAC.integrated, features = features, assay = 'RNA', cols = c('White', 'Blue')) + RotatedAxis() + xlab('Marker Genes')+ ylab('Clusters')

annotations <- c('T', 'NK', 'DC1', 'MYE', 'UNK', 'DC3', 'UNK', 'MYE', 'DC3', 'DC3', 'PERI+FIB', 'MAST', 'MYE', 'PLASMA', 'DC1', 'B', 'UNK', 'PERI', 'DC2', 'DC1', 'MYE+TNK', 'ACINAR', 'DC1', 'ENDO', 'PDC', 'DC1', 'PLAS+MAST')

Idents(PDAC.integrated) <- PDAC.integrated$seurat_clusters
names(annotations) <- levels(PDAC.integrated)
PDAC.integrated <- RenameIdents(PDAC.integrated, annotations)
PDAC.integrated@meta.data$seurat_clusters <- PDAC.integrated@active.ident

PDAC.integrated.umap <- DimPlot(PDAC.integrated, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.integrated.umap 

## Rename patients 

PDAC.integrated$orig.ident <- as.factor(PDAC.integrated$orig.ident)

levels(PDAC.integrated@meta.data$orig.ident) <- list(P1 = 'PDAC155698.1', P2 = 'PDAC155698.2', P3 = 'PDAC155698.3', P4 = 'PDAC155698.4', P5 = 'PDAC155698.5', P6 = 'PDAC155698.6', P7 = 'PDAC155698.7', P8 = 'PDAC155698.8', P9 = 'PDAC155698.9', P10='PDAC155698.10', P11A = 'PDAC155698.11A', P11B = 'PDAC155698.11B', P12 = 'PDAC155698.12', P13 = 'PDAC155698.13', P15 = 'PDAC155698.15', P16 = 'PDAC155698.16')

PDAC.integrated$patient <- PDAC.integrated$orig.ident
PDAC.integrated$orig.ident <- NULL

```



Subcluster immune cells 

```{r}

Idents(PDAC.integrated) <- PDAC.integrated$seurat_clusters

PDAC.integrated.metadata <- PDAC.integrated@meta.data

PDAC.integrated.metadata$seurat_clusters <- as.character(PDAC.integrated.metadata$seurat_clusters)

## Update clusters 0 and 1 (T and NK cells)

for (i in 1:nrow(PDAC.integrated.metadata)) {
  if (PDAC.integrated.metadata$RNA_snn_res.0.5[i] == 0 || PDAC.integrated.metadata$RNA_snn_res.0.5[i] == 1) {
    PDAC.integrated.metadata$seurat_clusters[i] <- PDAC.integrated.metadata$sub_cluster[i]
  }
}

PDAC.integrated@meta.data <- PDAC.integrated.metadata

PDAC.integrated.umap <- DimPlot(PDAC.integrated, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.integrated.umap 


## Update cluster 20 (Myeloid and T cells)

for (i in 1:nrow(PDAC.integrated.metadata)) {
  if (PDAC.integrated.metadata$RNA_snn_res.0.5[i] == 20) {
    PDAC.integrated.metadata$seurat_clusters[i] <- PDAC.integrated.metadata$sub_cluster[i]
  }
  if (PDAC.integrated.metadata$seurat_clusters[i] == 'Other') {
    PDAC.integrated.metadata$seurat_clusters[i] <- "MYE"
  }
}

PDAC.integrated@meta.data <- PDAC.integrated.metadata

PDAC.integrated.umap <- DimPlot(PDAC.integrated, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.integrated.umap 


## Subcluster ductal cells 

PDAC.DC <- subset(PDAC.integrated, idents=c("DC1", "DC2", "DC3"))
PDAC.DC@meta.data$orig.cluster <- PDAC.DC@meta.data$seurat_clusters

PDAC.DC <- clustering(PDAC.DC, resolution=0.3)

features <- c('TFF1', 'TFF2', 'LYZ', 'VSIG2', 'CEACAM6', 'APOA1', 'FABP1', 'ANPEP', 'ADIRF', 'DNASE1', 'CTRB2', 'CFTR', 'SOD3', 'COL18A1', 'VCAM1', 'CRISP3', 'TCN1', 'MMP1', 'SPINK1', 'LGALS2')

DotPlot(PDAC.DC, features = features, assay = 'RNA', cols = c('White', 'Blue')) + RotatedAxis() + xlab('Marker Genes')+ ylab('Clusters')

annotations <- c("Ductal cells 1", "Ductal cells 1", "Ductal cells 1", "Ductal cells 4", "Ductal cells 4", "Ductal cells 1", "Ductal cells 1", "Ductal cells 1", "Ductal cells 1", "Ductal cells 1", "Ductal cells 2", "Ductal cells 1", "Ductal cells 3", "Ductal cells 1")

Idents(PDAC.DC) <- PDAC.DC$seurat_clusters
names(annotations) <- levels(PDAC.DC)
PDAC.DC <- RenameIdents(PDAC.DC, annotations)
PDAC.DC@meta.data$seurat_clusters <- PDAC.DC@active.ident

PDAC.DC.umap <- DimPlot(PDAC.DC, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.DC.umap

PDAC.integrated$sub_cluster1 <- as.character(Idents(PDAC.integrated))
PDAC.integrated$sub_cluster1[Cells(PDAC.DC)] <- paste(Idents(PDAC.DC))
DimPlot(PDAC.integrated, group.by="sub_cluster1", label = TRUE)


## Subcluster perivascular cells and fibroblasts

PDAC.peri.fib <- subset(PDAC.integrated, idents=c("PERI", "PERI+FIB"))
PDAC.peri.fib@meta.data$orig.cluster <- PDAC.peri.fib@meta.data$seurat_clusters

PDAC.peri.fib <- clustering(PDAC.peri.fib, resolution=0.5)

features <- c('IGFBP7', 'ACTA2', 'RGS5', 'COL1A1', 'COL3A1', 'LUM', 'DCN', 'IL6', 'PDGFRA', 'CXCL12', 'MMP11', 'MYL9')

DotPlot(PDAC.peri.fib, features = features, assay = 'RNA', cols = c('White', 'Blue')) + RotatedAxis() + xlab('Marker Genes')+ ylab('Clusters')

annotations <- c("Fibroblasts", "Fibroblasts", "Fibroblasts", "Perivascular cells", "Fibroblasts", "Fibroblasts", "Perivascular cells")

Idents(PDAC.peri.fib) <- PDAC.peri.fib$seurat_clusters
names(annotations) <- levels(PDAC.peri.fib)
PDAC.peri.fib <- RenameIdents(PDAC.peri.fib, annotations)
PDAC.peri.fib@meta.data$seurat_clusters <- PDAC.peri.fib@active.ident

PDAC.peri.fib.umap <- DimPlot(PDAC.peri.fib, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.peri.fib.umap

Idents(PDAC.integrated) <- PDAC.integrated$sub_cluster1
PDAC.integrated$sub_cluster2 <- as.character(Idents(PDAC.integrated))
PDAC.integrated$sub_cluster2[Cells(PDAC.peri.fib)] <- paste(Idents(PDAC.peri.fib))
DimPlot(PDAC.integrated, group.by="sub_cluster2", label = TRUE)


## Subcluser plasma and mast cells

PDAC.plas.mast <- subset(PDAC.integrated, idents=c("PLASMA", "PLAS+MAST", "MAST"))
PDAC.plas.mast@meta.data$orig.cluster <- PDAC.plas.mast@meta.data$seurat_clusters

PDAC.plas.mast <- clustering(PDAC.plas.mast, resolution=1.0)

features <- c('IGJ', 'IGLL5', 'CLU', 'TPSAB1', 'CPA3')

DotPlot(PDAC.plas.mast, features = features, assay = 'RNA', cols = c('White', 'Blue')) + RotatedAxis() + xlab('Marker Genes')+ ylab('Clusters')

annotations <- c("Mast cells", "Plasma cells", "Mast cells", "Mast cells", "Plasma cells", "Plasma cells", "Plasma cells", "Mast cells", "Plasma cells", "Mast cells", "Plasma cells")

Idents(PDAC.plas.mast) <- PDAC.plas.mast$seurat_clusters
names(annotations) <- levels(PDAC.plas.mast)
PDAC.plas.mast <- RenameIdents(PDAC.plas.mast, annotations)
PDAC.plas.mast@meta.data$seurat_clusters <- PDAC.plas.mast@active.ident

PDAC.plas.mast.umap <- DimPlot(PDAC.plas.mast, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.plas.mast.umap

Idents(PDAC.integrated) <- PDAC.integrated$sub_cluster2
PDAC.integrated$sub_cluster3 <- as.character(Idents(PDAC.integrated))
PDAC.integrated$sub_cluster3[Cells(PDAC.plas.mast)] <- paste(Idents(PDAC.plas.mast))
DimPlot(PDAC.integrated, group.by="sub_cluster3", label = TRUE)


## Subcluster myeloid and unknown cells 

PDAC.unk <- subset(PDAC.integrated, idents=c("UNK", "MYE"))
PDAC.unk@meta.data$orig.cluster <- PDAC.unk@meta.data$seurat_clusters

PDAC.unk <- clustering(PDAC.unk, resolution=0.5)

features <- c('LAMP3', 'CCL22', 'TFF1', 'KRT18', 'KRT19', 'KRT8', 'CLU', 'MMP7', 'SPP1', 'REG1A', 'CTRB2', 'PRSS1', 'DCN', 'LUM', 'CPA3', 'TPSAB1', 'CDH5', 'VWF', 'PLVAP', 'IRF7', 'RGS5', 'PDGFRB', 'CD3E', 'NCAM1', 'NKG7', 'CD3D', 'CD14', 'HLA-DRA', 'GZMB', 'ITGAX', 'ITGAM', 'APOE', 'LYZ', 'IGJ', 'CD79A', 'MS4A1')

DotPlot(PDAC.unk, features = features, assay = 'RNA', cols = c('White', 'Blue')) + RotatedAxis() + xlab('Marker Genes')+ ylab('Clusters')

annotations <- c("Granulocytes", "Macrophages", "Epithelial cells", "Macrophages", "Macrophages", "Macrophages", "Granulocytes", "Macrophages", "Granulocytes", "Macrophages", "Granulocytes", "Macrophages", "Epithelial cells", "Macrophages")

Idents(PDAC.unk) <- PDAC.unk$seurat_clusters
names(annotations) <- levels(PDAC.unk)
PDAC.unk <- RenameIdents(PDAC.unk, annotations)
PDAC.unk@meta.data$seurat_clusters <- PDAC.unk@active.ident

PDAC.unk.umap <- DimPlot(PDAC.unk, reduction="umap", label=TRUE, pt.size=0.1)
PDAC.unk.umap

Idents(PDAC.integrated) <- PDAC.integrated$sub_cluster3
PDAC.integrated$sub_cluster4 <- as.character(Idents(PDAC.integrated))
PDAC.integrated$sub_cluster4[Cells(PDAC.unk)] <- paste(Idents(PDAC.unk))
DimPlot(PDAC.integrated, group.by="sub_cluster4", label = TRUE)


## Rename all idents

Idents(PDAC.integrated) <- PDAC.integrated$sub_cluster4

PDAC.integrated <- RenameIdents(object = PDAC.integrated,
                                'ACINAR' = 'Acinar cells',
                                'B' = 'B cells', 
                                'PDC' = 'pDCs',
                                'ENDO' = 'Endothelial cells')

PDAC.integrated@meta.data$seurat_clusters <- PDAC.integrated@active.ident

PDAC.integrated@meta.data$sub_cluster <- NULL
PDAC.integrated@meta.data$sub_cluster1 <- NULL
PDAC.integrated@meta.data$sub_cluster2 <- NULL
PDAC.integrated@meta.data$sub_cluster3 <- NULL
PDAC.integrated@meta.data$sub_cluster4 <- NULL

Idents(PDAC.integrated) <- PDAC.integrated$seurat_clusters
PDAC.integrated.umap <- DimPlot(PDAC.integrated, group.by="seurat_clusters", label=TRUE)
PDAC.integrated.umap

save(PDAC.integrated, file="C:\\Users\\Ryan\\Documents\\Nie Research Project\\GSE 155698\\PDAC.integrated")

```





